How did local health infrastructure and socio-political factors within different states and counties in the United States affect the disparities in COVID-19 outcomes, and what lessons can be learned for more targeted public health preparedness and response strategies in future pandemics?

### 2024-02-12 outputs/paper/paper.qmd
fixed paper and folders

---
title: "How did local health infrastructure and socio-political factors within different states and counties in the United States affect the disparities in COVID-19 outcomes, and what lessons can be learned for more targeted public health preparedness and response strategies in future pandemics?"
author: 
  - Adrian Ly
  - Sakhil Goel
  - Hannah Yu
thanks: "Code and data are available at: https://github.com/hannahyu07/US-Covid-Analysis.git"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
toc: true
prefer-html: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false
#### Workspace setup ####
library(tidyverse)
library(sf)
library(readxl)
library(knitr)
library(janitor)
library(lubridate)
library(dplyr)
library(data.table)
library(RColorBrewer)
library(ggpubr)
library(scales)
library(here)
library(readxl)
library(kableExtra)

@@

#ghs_data <- read_excel(here::here("inputs/data/GHS_index.xlsx"))
#world_map <- st_read(here::here("inputs/data/world/World_Countries__Generalized_.shp"))
#life_exp <- read.csv(here::here("inputs/data/life_table.csv"), stringsAsFactors = FALSE)
#vote_share <- read.csv(here::here("inputs/data/Vote_share_2020_data.csv"), stringsAsFactors = FALSE)
ghs_data <- read_excel("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/GHS_index.xlsx")
world_map <- st_read("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/world/World_Countries__Generalized_.shp")
life_exp <- read.csv("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/life_table.csv", stringsAsFactors = FALSE)
vote_share <- read.csv("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/Vote_share_2020_data.csv", stringsAsFactors = FALSE)
deaths <- read.csv("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/covid-19-death-rates.csv", stringsAsFactors = FALSE)

@@

`R` [@citeR] was the language and environment used for the bulk of this analysis, alongside `tidyverse` [@citeTidyVerse], `sf` [@citeSF], `readxl` [@citeReadXL], `knitr` [@citeKnitR], `janitor` [@citeJanitor], `lubridate` [@citeLubridate], `dplyr` [@citeDplyr], `data.table` [@citeDataTable], `RColorBrewer` [@citeRColorBrewer], `ggpubr` [@citeggpubr], `ggplot2` [@citeggplot], `here` [@citeHere], and `scales` [@citeScales].

@@

```{r}
#| echo: false
#| eval: true
#| label: tbl- deaths
#| tbl-cap: Top 10 States with Highest Death Rates from COVID-19 (per 100,000 people)
#| warning: false
deaths <- read.csv("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/covid-19-death-rates.csv", stringsAsFactors = FALSE)

@@

# Print the table using kableExtra
kable(table, caption = "Top 10 States with Highest Death Rates from COVID-19 (per 100,000 people)", align = "c", format = "html") %>%
  kable_styling(full_width = FALSE)
```

@@

## Weaknesses and next steps

>>>>

---
title: "How did local health infrastructure and socio-political factors within different states and counties in the United States affect the disparities in COVID-19 outcomes, and what lessons can be learned for more targeted public health preparedness and response strategies in future pandemics?"
author: 
  - Adrian Ly
  - Sakhil Goel
  - Hannah Yu
thanks: "Code and data are available at: https://github.com/hannahyu07/US-Covid-Analysis.git"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

#### Workspace setup ####
library(tidyverse)
library(sf)
library(readxl)
library(knitr)
library(janitor)
library(lubridate)
library(dplyr)
library(data.table)
library(RColorBrewer)
library(ggpubr)
library(scales)
library(here)
library(kableExtra)

@@

ghs_data <- read_excel(here::here("inputs/data/GHS_index.xlsx"))
world_map <- st_read(here::here("inputs/data/world/World_Countries__Generalized_.shp"))
life_exp <- read.csv(here::here("inputs/data/life_table.csv"), stringsAsFactors = FALSE)
vote_share <- read.csv(here::here("inputs/data/Vote_share_2020_data.csv"), stringsAsFactors = FALSE)
deaths <- read.csv(here::here("inputs/data/covid-19-death-rates.csv"), stringsAsFactors = FALSE)

@@

`R` [@citeR] was the language and environment used for the bulk of this analysis, alongside `tidyverse` [@citeTidyVerse], `sf` [@citeSF], `readxl` [@citeReadXL], `knitr` [@citeKnitR], `janitor` [@citeJanitor], `lubridate` [@citeLubridate], `dplyr` [@citeDplyr], `data.table` [@citeDataTable], `RColorBrewer` [@citeRColorBrewer], `ggpubr` [@citeggpubr], `ggplot2` [@citeggplot], `here` [@citeHere], `kableExtra` [@citeKableExtra], and `scales` [@citeScales].

@@

```{r}
#| echo: false
#| eval: true
#| label: tbl-deaths
#| tbl-cap: Top 10 States with Highest Death Rates from COVID-19 (per 100,000 people)
#| warning: false

@@

# Print the table using kableExtra
kable(table, caption = "Top 10 States with Highest Death Rates from COVID-19 (per 100,000 people)", align = "c") %>%
  kable_styling(full_width = FALSE)

@@

## Weaknesses and next stepsa
